---
title: "Get started with ggplotpq"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with ggplotpq}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5
)
```

## What ggplotpq is for

`ggplotpq` is the **`ggplot2` visualisation layer** of the
[pqverse](https://github.com/adrientaudiere). It owns pure `ggplot2` helpers
(geoms, themes, scales, colour/palette management) and single-`phyloseq`
`ggplot` wrappers that previously lived in
[`MiscMetabar`](https://adrientaudiere.github.io/MiscMetabar/) or `comparpq`.

It is deliberately **not** for cross-package multi-`phyloseq` comparators
(those stay in `comparpq`), nor for machine-learning / networks (those stay in
`netaipq`). Genuinely new analysis methods belong in `netaipq`; `ggplotpq` may
render their plots, but does not host the analysis.

## Installation

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("adrientaudiere/ggplotpq")
```

```{r setup, message = FALSE}
library(ggplotpq)
library(ggplot2)
data(data_fungi_mini, package = "MiscMetabar")
```

## Plotting helpers for phyloseq objects

`plot_sample_depth_pq()` summarises sequencing depth across samples — a quick
first diagnostic before any rarefaction or normalisation.

```{r depth}
plot_sample_depth_pq(data_fungi_mini)
```

`plot_taxa_heatmap_pq()` shows the most abundant taxa across samples, at the
taxonomic rank of your choice.

```{r heatmap}
plot_taxa_heatmap_pq(data_fungi_mini, n_top = 15, taxa_rank = "Family")
```

`gg_bubbles_pq()` lays out taxa as bubbles sized by abundance and coloured by a
taxonomic rank.

```{r bubbles}
gg_bubbles_pq(physeq = data_fungi_mini, rank_color = "Class")
```

## Themes, palettes, and scales

`ggplotpq` ships reusable `ggplot2` building blocks that work on any plot, not
only `phyloseq` ones. `theme_pq_minimal()` is a clean, publication-oriented
theme:

```{r theme}
ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  theme_pq_minimal()
```

The `scale_*_pq_discrete()` family gives curated discrete palettes:

```{r scale}
ggplot(mtcars, aes(wt, mpg, color = factor(cyl))) +
  geom_point(size = 2) +
  scale_color_pq_discrete("Picabia") +
  theme_pq_minimal()
```

The `idest` family (`theme_idest()`, `idest_pal()`, `scale_color_idest_*()`,
`scale_fill_idest_*()`) provides a coordinated branded look; see the
[function reference](https://adrientaudiere.github.io/ggplotpq/reference/index.html) for the full set.

## Where to go next

- The [function reference](https://adrientaudiere.github.io/ggplotpq/reference/index.html) is organised by origin
  (migrated from `MiscMetabar`, from `comparpq`, from `taxinfo`) and by new
  visualisation helpers.
- Because every function returns a standard `ggplot` object, you can freely
  combine `ggplotpq` helpers with each other and with the wider `ggplot2`
  ecosystem (`patchwork`, custom scales, and so on).
